A Plasma Torus Around a Young Low-Mass Star

### 2025-06-10 paper/ms.tex
ran clean.html

\begin{abstract}
  A small fraction of red dwarfs younger than $\sim$100 million years
  show structured, periodic optical light curves suggestive of
  transiting opaque material that corotates with the star.  However, the
  composition, origin, and even the existence of this material are
  uncertain. The main alternative hypothesis is that these complex
  periodic variables (CPVs) are explained by complex distributions of
  bright or dark regions on the stellar surfaces.  Here, we present
  time-series spectroscopy and photometry of a rapidly-rotating
  ($P$=3.9\,hr) CPV, TIC\,141146667.  The spectra show sinusoidal
  time-varying H$\alpha$ emission at twice to four times the star's
  equatorial velocity, providing direct evidence for cool
  ($\lesssim$10$^4$\,K) plasma clumps trapped in corotation around a
  CPV.  These data support the idea that young, rapidly-rotating M
  dwarfs can sustain warped \added{tori of cool plasma, similar to
  other rapidly-rotating magnetic stars.}
  \added{Outstanding questions include whether dust clumps 
  in these plasma tori explain CPV light curves, and
  whether the tori originate from the star or are fed by external
  sources.}
  Rough estimates suggest $\gtrsim$10\% of M dwarfs host
  similar structures during their early lives.
\end{abstract}

@@

Stars with masses below about half that of the Sun, M dwarfs, are the
only type of star to offer near-term prospects for detecting the
atmospheres of rocky exoplanets with surface water.  Community
investment with JWST is proceeding accordingly
\citep[][]{Redfield2024,TRAPPIST1JWSTCommunityInitiative2024}.  The
question of how M dwarfs influence their planets---especially the
retention of their atmospheres---has correspondingly grown in
importance.  Previous work has established that most M dwarfs host
close-in planets \citep{Dressing2015} that on average are subject to
long disk lifetimes \citep{Ribas2015}, intense UV radiation
\citep{France2016}, and frequent stellar flares \citep{Feinstein2020}.
However, despite extensive work in these areas, the plasma and
magnetospheric environments that bathe young, close-in exoplanets
remain challenging to quantify.  Understanding these environments is
crucial because they directly impact atmospheric retention and\added{,
ultimately,} habitability.

@@

\begin{figure*}[!t]
  \centering
  \includegraphics[width=0.925\textwidth]{f2.pdf}
  \vspace{-0.2cm}
  \caption{{\bf Photometric evolution of TIC\,141146667 around the Keck/HIRES
  observation (green bar).}  {\bf a,} TESS simple aperture photometry.
  Data gaps were caused by stray light from the Earth (BTJD 3356-3358.5) and
  Moon (BTJD 3359.5-3360.5).  Raw two minute data are in gray; black
  time-averages to ten minute sampling.  {\bf b-c,} Folded TESS light
  curve before and after spectroscopy.  Black now phase-averages to
  100 points per 3.93\ hour cycle.  During BTJD 3352-3356, a state
  switch occurred near BTJD 3353, and the dip at $\phi$$\approx$0.8
  \added{became less pronounced}.  While the \added{dip spanning
  $\phi$=0.6-1.1} was present both before and after the HIRES
  sequence, its photometric shape evolved during the data gap.  The
  orange bar denotes times of spectroscopic transits for the inner two
  H$\alpha$ clumps observed with HIRES (see Figure~\ref{fig:spec}).}
  \label{fig:fulllc}
\end{figure*}

\begin{figure*}[!t]
  \centering
  \includegraphics[width=0.99\textwidth]{f3.pdf}
  \vspace{-0.3cm}
  \captionsetup{labelformat=moviefmt,labelsep=colon}
	\caption{\textbf{Emission from plasma clumps orbiting
TIC\,141146667.}
  The online movie,
  \href{https://lgbouma.com/movies/TIC141146667_sixpanel.mp4}{{\bf
  available here}},
  shows the spectral evolution over five hours.
  {\bf a,} TESS light curve from 5 February 2024 to 26
  February 2024 folded on the 3.93\ hour period.  Black
  points are phase-averaged; gray are the raw data.
  {\bf b,} Keck/HIRES H$\alpha$ spectra from 17 February 2024.
  The continuum is set to unity, and the darkest color is at twice the
  continuum to accentuate emission outside the line core ($|v/v_{\rm
  eq}|>1$, for the equatorial velocity $v_{\rm eq}$=130\,\kms).  While
  emission within the line core comes from the star's chromosphere, the
  sinusoidal emission features are most readily described by a warped
  plasma torus.
  {\bf c,} Individual epochs of Panel b, visible in the online movie.
  The dotted line shows the time-averaged spectrum, $f_{\langle t
  \rangle}$.
  {\bf d,} As in Panel a, but overplotting the
  median-normalized H$\alpha$ light curve at $|v/v_{\rm eq}|<1$.
  {\bf e,} As in Panel b, after subtracting $f_{\langle t
  \rangle}$.  The line core shows H$\alpha$ excesses and decrements
  advancing from the blue to red wings.
  The asymmetric color stretch mirrors the dynamic range of the data.
  {\bf f,} Individual epochs of Panel e, visible in the online
  movie.  \added{A separate version with best-fit sinusoids from
  Appendix~\ref{subsec:halpha} overplotted is also
  \href{https://lgbouma.com/movies/TIC141146667_sixpanel_sinusoids.mp4}{{\bf
  available}}.}}
  \label{fig:spec}
\end{figure*}

@@

We observed TIC\,141146667 ($V$=16.2) for five hours on
17~February~2024 using the High Resolution Echelle Spectrometer
(HIRES; \citealt{vogt_hires_1994}) on the 10\,m Keck I telescope.  The
observations spanned the second half of the night, from 10:47 to 16:13
(UT).  \added{We opted for a fixed 15 minute cadence, and} used the C2
decker (0$\farcs$86$\times$14$\farcs$0) in the red configuration,
yielding a resolution $R$$\approx$45{,}000 ($\delta
v$$\approx$6.7\,\kms).  Strong winds contributed to 1$''$-1\farcs5
seeing, but conditions were otherwise favorable.  We reduced the
echelleogram using the standard techniques of the California Planet
Survey \citep{Howard2010}.

TESS observed TIC\,141146667 ($T$=13.3) for six non-contiguous months
spanning 2019-2024.  TESS acquired these observations at two-minute
cadence during Sectors 41, 48, and 75.   The 30-minute data during
Sectors 14, 15, and 21 smeared sharp features over the star's
3.93~hour period (see \citealt{Gunther2022}).  The movie in
Figure~\ref{fig:lc} shows the two-minute data: like other CPVs,
TIC\,141146667 maintains a fixed period over timescales of years while
its detailed photometric morphology evolves.  The nearest known star,
TIC\,141146666 ($T$=14.5), is 25$''$ from TIC\,141146667 and is
photometrically \added{quiet}; crowding is not a concern.

@@

Figure~\ref{fig:spec} shows the TESS and HIRES data from February
2024.  The spectra \added{(Panels b,c)} show emission in H$\alpha$
beyond the star's equatorial velocity, $v_{\rm eq}$, of 130\,\kms.
There are at least two distinct emission components, 180$^\circ$ apart
in phase.  The inner component at lower velocities has clearer
sinusoidal behaviour in time and is double-peaked, with peak
semi-amplitudes of 2.07\,$v_{\rm eq}$ and 2.88\,$v_{\rm eq}$ (see
Appendix~\ref{subsec:halpha}).  There is significant non-periodic
variability in the emissivity of this double-peaked component: the
flux excess begins with an amplitude 70\% above the continuum, and
diminishes to 10\% by sunrise.  The higher-velocity component
180$^\circ$ opposite in phase is detected only from $\phi$=0.2-1.0.
From $\phi$=0.2-0.5, this outer component appears connected to the
star in velocity space.  While its peak semi-amplitude of
3.88\,$v_{\rm eq}$ is achieved at both $\phi$=0.25 and $\phi$=0.75,
its amplitude similarly decreases from a 60\% excess to a 10\% excess
over the observation sequence.  In Appendix~\ref{subsec:halpha}, we
measured the period for all three emission components to be consistent
with the photometric \periodhr\ hour period, to within two minutes for
the \added{inner} component.

These sinusoidal emission features require circumstellar clumps of
partially-ionized hydrogen to corotate with the star.  Based on the
observed sinusoidal periods and velocities, this \added{material is
not moving on a Keplerian orbit}; it is \added{magnetically forced to
co-rotate with the star}.  The velocity semi-amplitude of the
sinusoids gives the mean distance of each clump from the stellar
\added{rotation axis}: 2.07\,$R_\star$ and 2.88\,$R_\star$ for the
inner clumps, and 3.88\,$R_\star$ for the outer clump.   These clumps
transit in front of the star when passing from negative to positive
velocity.  The transits of the two inner clumps last $\approx$22\% of
each cycle, from $\phi$=-0.1 to $\phi$=+0.12.  This spectroscopic
transit coincides with the latter half of the complex eclipse feature
in the TESS data \added{(see Figure~\ref{fig:fulllc} and Panels a and
b of Figure~\ref{fig:spec})}.

The H$\alpha$ line core\added{, visible in Panels e and f of
Figure~\ref{fig:spec}, }is more complex.  At $|\Delta v / v_{\rm
eq}|<1$, most observed H$\alpha$ photons come from the star's
chromosphere; \added{variability in the stellar emission line can be
caused by circumstellar material, bright regions, or dark regions
crossing the star's surface}.  Figure~\ref{fig:spec}e suggests that
\added{all three effects occur in TIC\,141146667}.  For instance, from
$\phi$=0-0.3, double-peaked emission is visible both on and off-limb;
this feature is circumstellar in origin.  However, the emission
feature that crosses the star from $\phi$=0.4-0.9 is brighter than the
circumstellar clumps, and its slow speed instead suggests it is
associated with a bright region on the star's surface.  Similarly, from
$\phi$=0.6-1.15 a 20\% deep absorption feature slowly crosses the
H$\alpha$ line profile.  This feature could be either a
chromospherically dark region (e.g.~a spot group), or an azimuthally
extended component of the circumstellar material.  The origin of the
other bright and dark streaks passing across the line core \added{in
Figure~\ref{fig:spec}e }is similarly ambiguous.

\added{Figure~\ref{fig:spec}d shows} a final exercise to quantify the
behavior of the line core, by summing the \added{H$\alpha$} flux at
$|\Delta v / v_{\rm eq}|<1$.  This panel shows that changes in the
line core flux \added{($f_{\rm H\alpha\ core}$)} correlate with the
broadband variability throughout most of the light curve, except near
$\phi$$\approx$0.5, corresponding to the transit of the 3.9\,$R_\star$
clump and the occultation of the lower-velocity clump.
\added{Spectroscopic transits may therefore not always be associated
with photometric transits.}

@@

Our Keck/HIRES observations are the first reported time-series spectra
of a CPV, and they demonstrate that corotating circumstellar plasma
clumps exist around at least one such star.  More specifically, the
spectra require plasma with a significant population of hydrogen in
the $n$=3 excited state, \added{with }minimal evidence for
higher-order excitations.  Most of this H$\alpha$ emission originates
in ``clumps'' with size comparable to the star; radial ``spokes'' or
azimuthally extended ``arcs'' for the emitting material are ruled out
by the $\approx$30\,\kms\ H$\alpha$ velocity dispersion (see
Appendix~\ref{subsec:halpha}).  The H$\alpha$ line luminosity suggests
characteristic number densities and masses for the gaseous component
of these clumps of $n_{\rm H} \sim 10^{11}$\,cm$^{-3}$ and $M_{\rm
gas} \sim 10^{17}$\,g (see Appendix~\ref{subsec:gas}).  Dust is
independently constrained to have a total mass $M_{\rm dust} <
10^{17}\,{\rm g}$ based on the lack of a WISE infrared excess; if one
assumes that the opacity in the TESS flux dips comes from dust, a
lower limit $M_{\rm dust} > 10^{15}\,{\rm g}$ follows (see
Appendix~\ref{subsec:dust}).

@@

While the H$\alpha$ emission mostly comes from clumps, the broadband
flux dip in Figure~\ref{fig:spec} spans roughly half of each cycle.
In CPVs more generally, individual flux dips last 5-50\% of each
cycle\added{, and there are often multiple dips per cycle
}\citep{Bouma2024}.  This implies that azimuthally-distributed
material may be the norm for CPVs; a warped torus may be a more
accurate picture than a clump.

\citet{Townsend2005} outlined the physics of how rapidly-rotating
stars with strong magnetic fields can support plasma tori
\added{\citep[see
also][]{Nakajima1985,Ferreira2000,Petit2013,Daley-Yates2024}}.  When
the magnetospheric radii $R_{\rm m}$ of such stars exceeds their
Keplerian corotation radii $R_{\rm c}$, the effective potential\added{
  (gravitational plus centrifugal)} experienced by \added{plasma along
any given field line }has a local minimum outside $R_{\rm c}$, which
enables charged material to accumulate in a torus.  Warps can occur
when there is misalignment between the spin and magnetic axes.  In
general, the\added{ regions} in these centrifugally-supported
magnetospheres\added{ with the densest plasma accumulations} need
neither transit nor be opaque in broadband optical light.

@@

To summarize, the evidence for warped plasma tori in TIC\,141146667
and CPVs more broadly is that Figure~\ref{fig:spec} shows two plasma
clumps separated by 180$^\circ$ in phase; the warped-torus model
predicts both ``W'' shaped photometric eclipses and double-peaked
H$\alpha$ morphology at quadrature; and CPV dips last 5-50\% of each
cycle, indicating an azimuthally extended distribution of material.
\added{However, two challenges remain. First, the model predicts {\it
two} W-shaped eclipses per cycle.  Second, it predicts that the
antipodal clumps should lie at equal distances from the star.
Different clump opacities or non-dipolar magnetic fields might resolve
these discrepancies.  Overall, the evidence suggests a warped torus is
a good first approximation, but one that will need refinement in
future work.}

@@

The other CPV analog is the $\sigma$~Ori~E variables, a rare subset of
B stars with radiatively-driven winds that accumulate into warped
plasma tori \citep{Townsend2005}.  These tori tend to have dense
antipodal accumulations of plasma sculpted by tilted-dipole magnetic
fields; these clumps produce broadband optical variability through
bound-free scattering \citep{Townsend2005} and Thomson scattering
\citep{Berry2022}.  For $\sigma$~Ori~E and most of its analogs, the
result is simple light curves that resemble eclipsing binaries, and
time-dynamic H$\alpha$ spectra similar to Figure~\ref{fig:spec}. 
\added{The few known exceptions---HD~37776, HD~64740, and HD~176582---}have complex
light curves resembling CPVs \citep{Mikulasek2020,Bouma2024} and
spectropolarimetric measurements that suggest non-dipolar field
contributions \citep{Kochukhov2011,Shultz2018}.  This suggests that
the photometric complexity of CPVs may be related to magnetic fields
with highly multipolar contributions.  \added{While this suggestion is
consistent with the wide range of axisymmetric and non-axisymmetric
topologies seen in mid- and late-M dwarfs
(e.g.~\citealt{Donati2006,Kochukhov2017,Shulyak2019,Bellotti2024}),
future work is needed to verify whether CPVs indeed select for
complex, non-axisymmetric fields.}

@@

Spectra of magnetically-active, rapidly rotating stars with a wide
range of masses have been previously observed to exhibit sinusoidal
time-varying Balmer emission
\citep{Donati2000,Townsend2005b,Dunstone2006,Skelly2008}, similar to
Figure~\ref{fig:spec}.  No such stars were previously known to show
complex light curves \citep{Bouma2024}.   One interpretation for the
spectroscopic variability of these stars, and that of the analogous
transient \added{H$\alpha$ }absorption features
\citep{CollierCameron1989,CollierCameron1992,Cang2020}, comes from an
analogy to quiescent solar prominences, cool condensations of plasma
in the solar corona that can last days to weeks
\citep{VialEngvold2015}.  In the case of the Sun, these condensations
fall back to the photosphere because gravity is stronger than any
magnetic tension or centrifugal force capable of sustaining them.
However, it has been understood at least since work by
\citet{Donati2000} that these ``prominence systems'' can be
longitudinally extended, forming a trapped ring of plasma.  While the
issue of why these previously known spectroscopic variables do not
show complex photometric variability remains open, plausible
explanations include that they are not in the required transiting
geometry, or that they lack the necessary source of opacity.  What we
have in the case of TIC\,141146667 is a system that finally exhibits
both sets of phenomena.

@@

The material's composition -- either pure plasma or dusty plasma --
can be clarified by time-series infrared spectrophotometry.  While
observations of CPVs in the optical suggest chromaticity consistent
with dust \citep{Tanimoto2020,Gunther2022,Koen2023}, electron
scattering in a plasma transiting over a spotted star might
also produce chromatic features \citep{Rackham2018}.  \added{This
degeneracy is alleviated in the infrared, where the} composition and
size distribution of any dust that is present could be determined by
measuring the extinction curve for a sample of CPVs from 1-20\,$\mu$m.
Composition and size distributions similar to debris from rocky bodies
seen around white dwarfs \citep{Reach2009} would suggest an extrinsic
origin channel.  Compositions and sizes similar to the interstellar
medium would suggest that dust can condense out from M dwarf winds,
similar to processes that occur around evolved stars
\citep{Marigo2008}.

>>>>

\begin{abstract}
  A small fraction of red dwarfs younger than $\sim$100 million years
  show structured, periodic optical light curves suggestive of
  transiting opaque material that corotates with the star.  However, the
  composition, origin, and even the existence of this material are
  uncertain. The main alternative hypothesis is that these complex
  periodic variables (CPVs) are explained by complex distributions of
  bright or dark regions on the stellar surfaces.  Here, we present
  time-series spectroscopy and photometry of a rapidly-rotating
  ($P$=3.9\,hr) CPV, TIC\,141146667.  The spectra show sinusoidal
  time-varying H$\alpha$ emission at twice to four times the star's
  equatorial velocity, providing direct evidence for cool
  ($\lesssim$10$^4$\,K) plasma clumps trapped in corotation around a
  CPV.  These data support the idea that young, rapidly-rotating M
  dwarfs can sustain warped tori of cool plasma, similar to
  other rapidly-rotating magnetic stars.
  Outstanding questions include whether dust clumps 
  in these plasma tori explain CPV light curves, and
  whether the tori originate from the star or are fed by external
  sources.
  Rough estimates suggest $\gtrsim$10\% of M dwarfs host
  similar structures during their early lives.
\end{abstract}

@@

Stars with masses below about half that of the Sun, M dwarfs, are the
only type of star to offer near-term prospects for detecting the
atmospheres of rocky exoplanets with surface water.  Community
investment with JWST is proceeding accordingly
\citep[][]{Redfield2024,TRAPPIST1JWSTCommunityInitiative2024}.  The
question of how M dwarfs influence their planets---especially the
retention of their atmospheres---has correspondingly grown in
importance.  Previous work has established that most M dwarfs host
close-in planets \citep{Dressing2015} that on average are subject to
long disk lifetimes \citep{Ribas2015}, intense UV radiation
\citep{France2016}, and frequent stellar flares \citep{Feinstein2020}.
However, despite extensive work in these areas, the plasma and
magnetospheric environments that bathe young, close-in exoplanets
remain challenging to quantify.  Understanding these environments is
crucial because they directly impact atmospheric retention and,
ultimately, habitability.

@@

\begin{figure*}[!t]
  \centering
  \includegraphics[width=0.925\textwidth]{f2.pdf}
  \vspace{-0.2cm}
  \caption{{\bf Photometric evolution of TIC\,141146667 around the Keck/HIRES
  observation (green bar).}  {\bf a,} TESS simple aperture photometry.
  Data gaps were caused by stray light from the Earth (BTJD 3356-3358.5) and
  Moon (BTJD 3359.5-3360.5).  Raw two minute data are in gray; black
  time-averages to ten minute sampling.  {\bf b-c,} Folded TESS light
  curve before and after spectroscopy.  Black now phase-averages to
  100 points per 3.93\ hour cycle.  During BTJD 3352-3356, a state
  switch occurred near BTJD 3353, and the dip at $\phi$$\approx$0.8
  became less pronounced.  While the dip spanning
  $\phi$=0.6-1.1 was present both before and after the HIRES
  sequence, its photometric shape evolved during the data gap.  The
  orange bar denotes times of spectroscopic transits for the inner two
  H$\alpha$ clumps observed with HIRES (see Figure~\ref{fig:spec}).}
  \label{fig:fulllc}
\end{figure*}

\begin{figure*}[!t]
  \centering
  \includegraphics[width=0.99\textwidth]{f3.pdf}
  \vspace{-0.3cm}
  \captionsetup{labelformat=moviefmt,labelsep=colon}
	\caption{\textbf{Emission from plasma clumps orbiting
TIC\,141146667.}
  The online movie,
  \href{https://lgbouma.com/movies/TIC141146667_sixpanel.mp4}{{\bf
  available here}},
  shows the spectral evolution over five hours.
  {\bf a,} TESS light curve from 5 February 2024 to 26
  February 2024 folded on the 3.93\ hour period.  Black
  points are phase-averaged; gray are the raw data.
  {\bf b,} Keck/HIRES H$\alpha$ spectra from 17 February 2024.
  The continuum is set to unity, and the darkest color is at twice the
  continuum to accentuate emission outside the line core ($|v/v_{\rm
  eq}|>1$, for the equatorial velocity $v_{\rm eq}$=130\,\kms).  While
  emission within the line core comes from the star's chromosphere, the
  sinusoidal emission features are most readily described by a warped
  plasma torus.
  {\bf c,} Individual epochs of Panel b, visible in the online movie.
  The dotted line shows the time-averaged spectrum, $f_{\langle t
  \rangle}$.
  {\bf d,} As in Panel a, but overplotting the
  median-normalized H$\alpha$ light curve at $|v/v_{\rm eq}|<1$.
  {\bf e,} As in Panel b, after subtracting $f_{\langle t
  \rangle}$.  The line core shows H$\alpha$ excesses and decrements
  advancing from the blue to red wings.
  The asymmetric color stretch mirrors the dynamic range of the data.
  {\bf f,} Individual epochs of Panel e, visible in the online
  movie.  A separate version with best-fit sinusoids from
  Appendix~\ref{subsec:halpha} overplotted is also
  \href{https://lgbouma.com/movies/TIC141146667_sixpanel_sinusoids.mp4}{{\bf
  available}}.}
  \label{fig:spec}
\end{figure*}

@@

We observed TIC\,141146667 ($V$=16.2) for five hours on
17~February~2024 using the High Resolution Echelle Spectrometer
(HIRES; \citealt{vogt_hires_1994}) on the 10\,m Keck I telescope.  The
observations spanned the second half of the night, from 10:47 to 16:13
(UT).  We opted for a fixed 15 minute cadence, and used the C2
decker (0$\farcs$86$\times$14$\farcs$0) in the red configuration,
yielding a resolution $R$$\approx$45{,}000 ($\delta
v$$\approx$6.7\,\kms).  Strong winds contributed to 1$''$-1\farcs5
seeing, but conditions were otherwise favorable.  We reduced the
echelleogram using the standard techniques of the California Planet
Survey \citep{Howard2010}.

TESS observed TIC\,141146667 ($T$=13.3) for six non-contiguous months
spanning 2019-2024.  TESS acquired these observations at two-minute
cadence during Sectors 41, 48, and 75.   The 30-minute data during
Sectors 14, 15, and 21 smeared sharp features over the star's
3.93~hour period (see \citealt{Gunther2022}).  The movie in
Figure~\ref{fig:lc} shows the two-minute data: like other CPVs,
TIC\,141146667 maintains a fixed period over timescales of years while
its detailed photometric morphology evolves.  The nearest known star,
TIC\,141146666 ($T$=14.5), is 25$''$ from TIC\,141146667 and is
photometrically quiet; crowding is not a concern.

@@

Figure~\ref{fig:spec} shows the TESS and HIRES data from February
2024.  The spectra (Panels b,c) show emission in H$\alpha$
beyond the star's equatorial velocity, $v_{\rm eq}$, of 130\,\kms.
There are at least two distinct emission components, 180$^\circ$ apart
in phase.  The inner component at lower velocities has clearer
sinusoidal behaviour in time and is double-peaked, with peak
semi-amplitudes of 2.07\,$v_{\rm eq}$ and 2.88\,$v_{\rm eq}$ (see
Appendix~\ref{subsec:halpha}).  There is significant non-periodic
variability in the emissivity of this double-peaked component: the
flux excess begins with an amplitude 70\% above the continuum, and
diminishes to 10\% by sunrise.  The higher-velocity component
180$^\circ$ opposite in phase is detected only from $\phi$=0.2-1.0.
From $\phi$=0.2-0.5, this outer component appears connected to the
star in velocity space.  While its peak semi-amplitude of
3.88\,$v_{\rm eq}$ is achieved at both $\phi$=0.25 and $\phi$=0.75,
its amplitude similarly decreases from a 60\% excess to a 10\% excess
over the observation sequence.  In Appendix~\ref{subsec:halpha}, we
measured the period for all three emission components to be consistent
with the photometric \periodhr\ hour period, to within two minutes for
the inner component.

These sinusoidal emission features require circumstellar clumps of
partially-ionized hydrogen to corotate with the star.  Based on the
observed sinusoidal periods and velocities, this material is
not moving on a Keplerian orbit; it is magnetically forced to
co-rotate with the star.  The velocity semi-amplitude of the
sinusoids gives the mean distance of each clump from the stellar
rotation axis: 2.07\,$R_\star$ and 2.88\,$R_\star$ for the
inner clumps, and 3.88\,$R_\star$ for the outer clump.   These clumps
transit in front of the star when passing from negative to positive
velocity.  The transits of the two inner clumps last $\approx$22\% of
each cycle, from $\phi$=-0.1 to $\phi$=+0.12.  This spectroscopic
transit coincides with the latter half of the complex eclipse feature
in the TESS data (see Figure~\ref{fig:fulllc} and Panels a and
b of Figure~\ref{fig:spec}).

The H$\alpha$ line core, visible in Panels e and f of
Figure~\ref{fig:spec}, is more complex.  At $|\Delta v / v_{\rm
eq}|<1$, most observed H$\alpha$ photons come from the star's
chromosphere; variability in the stellar emission line can be
caused by circumstellar material, bright regions, or dark regions
crossing the star's surface.  Figure~\ref{fig:spec}e suggests that
all three effects occur in TIC\,141146667.  For instance, from
$\phi$=0-0.3, double-peaked emission is visible both on and off-limb;
this feature is circumstellar in origin.  However, the emission
feature that crosses the star from $\phi$=0.4-0.9 is brighter than the
circumstellar clumps, and its slow speed instead suggests it is
associated with a bright region on the star's surface.  Similarly, from
$\phi$=0.6-1.15 a 20\% deep absorption feature slowly crosses the
H$\alpha$ line profile.  This feature could be either a
chromospherically dark region (e.g.~a spot group), or an azimuthally
extended component of the circumstellar material.  The origin of the
other bright and dark streaks passing across the line core in
Figure~\ref{fig:spec}e is similarly ambiguous.

Figure~\ref{fig:spec}d shows a final exercise to quantify the
behavior of the line core, by summing the H$\alpha$ flux at
$|\Delta v / v_{\rm eq}|<1$.  This panel shows that changes in the
line core flux ($f_{\rm H\alpha\ core}$) correlate with the
broadband variability throughout most of the light curve, except near
$\phi$$\approx$0.5, corresponding to the transit of the 3.9\,$R_\star$
clump and the occultation of the lower-velocity clump.
Spectroscopic transits may therefore not always be associated
with photometric transits.

@@

Our Keck/HIRES observations are the first reported time-series spectra
of a CPV, and they demonstrate that corotating circumstellar plasma
clumps exist around at least one such star.  More specifically, the
spectra require plasma with a significant population of hydrogen in
the $n$=3 excited state, with minimal evidence for
higher-order excitations.  Most of this H$\alpha$ emission originates
in ``clumps'' with size comparable to the star; radial ``spokes'' or
azimuthally extended ``arcs'' for the emitting material are ruled out
by the $\approx$30\,\kms\ H$\alpha$ velocity dispersion (see
Appendix~\ref{subsec:halpha}).  The H$\alpha$ line luminosity suggests
characteristic number densities and masses for the gaseous component
of these clumps of $n_{\rm H} \sim 10^{11}$\,cm$^{-3}$ and $M_{\rm
gas} \sim 10^{17}$\,g (see Appendix~\ref{subsec:gas}).  Dust is
independently constrained to have a total mass $M_{\rm dust} <
10^{17}\,{\rm g}$ based on the lack of a WISE infrared excess; if one
assumes that the opacity in the TESS flux dips comes from dust, a
lower limit $M_{\rm dust} > 10^{15}\,{\rm g}$ follows (see
Appendix~\ref{subsec:dust}).

@@

While the H$\alpha$ emission mostly comes from clumps, the broadband
flux dip in Figure~\ref{fig:spec} spans roughly half of each cycle.
In CPVs more generally, individual flux dips last 5-50\% of each
cycle, and there are often multiple dips per cycle
\citep{Bouma2024}.  This implies that azimuthally-distributed
material may be the norm for CPVs; a warped torus may be a more
accurate picture than a clump.

\citet{Townsend2005} outlined the physics of how rapidly-rotating
stars with strong magnetic fields can support plasma tori
\citep[see
also][]{Nakajima1985,Ferreira2000,Petit2013,Daley-Yates2024}.  When
the magnetospheric radii $R_{\rm m}$ of such stars exceeds their
Keplerian corotation radii $R_{\rm c}$, the effective potential
  (gravitational plus centrifugal) experienced by plasma along
any given field line has a local minimum outside $R_{\rm c}$, which
enables charged material to accumulate in a torus.  Warps can occur
when there is misalignment between the spin and magnetic axes.  In
general, the regions in these centrifugally-supported
magnetospheres with the densest plasma accumulations need
neither transit nor be opaque in broadband optical light.

@@

To summarize, the evidence for warped plasma tori in TIC\,141146667
and CPVs more broadly is that Figure~\ref{fig:spec} shows two plasma
clumps separated by 180$^\circ$ in phase; the warped-torus model
predicts both ``W'' shaped photometric eclipses and double-peaked
H$\alpha$ morphology at quadrature; and CPV dips last 5-50\% of each
cycle, indicating an azimuthally extended distribution of material.
However, two challenges remain. First, the model predicts {\it
two} W-shaped eclipses per cycle.  Second, it predicts that the
antipodal clumps should lie at equal distances from the star.
Different clump opacities or non-dipolar magnetic fields might resolve
these discrepancies.  Overall, the evidence suggests a warped torus is
a good first approximation, but one that will need refinement in
future work.

@@

The other CPV analog is the $\sigma$~Ori~E variables, a rare subset of
B stars with radiatively-driven winds that accumulate into warped
plasma tori \citep{Townsend2005}.  These tori tend to have dense
antipodal accumulations of plasma sculpted by tilted-dipole magnetic
fields; these clumps produce broadband optical variability through
bound-free scattering \citep{Townsend2005} and Thomson scattering
\citep{Berry2022}.  For $\sigma$~Ori~E and most of its analogs, the
result is simple light curves that resemble eclipsing binaries, and
time-dynamic H$\alpha$ spectra similar to Figure~\ref{fig:spec}. 
The few known exceptions---HD~37776, HD~64740, and HD~176582---have complex
light curves resembling CPVs \citep{Mikulasek2020,Bouma2024} and
spectropolarimetric measurements that may suggest non-dipolar field
contributions \citep{Kochukhov2011,Shultz2018}.  
The photometric complexity of CPVs may therefore be related to
magnetic fields with highly multipolar contributions.  While this
suggestion is consistent with the wide range of axisymmetric and
non-axisymmetric topologies seen in mid- and late-M dwarfs
(e.g.~\citealt{Donati2006,Kochukhov2017,Shulyak2019,Bellotti2024}),
future work is needed to verify whether CPVs indeed select for
complex, non-axisymmetric fields.

@@

Spectra of magnetically-active, rapidly rotating stars with a wide
range of masses have been previously observed to exhibit sinusoidal
time-varying Balmer emission
\citep{Donati2000,Townsend2005b,Dunstone2006,Skelly2008}, similar to
Figure~\ref{fig:spec}.  No such stars were previously known to show
complex light curves \citep{Bouma2024}.   One interpretation for the
spectroscopic variability of these stars, and that of the analogous
transient H$\alpha$ absorption features
\citep{CollierCameron1989,CollierCameron1992,Cang2020}, comes from an
analogy to quiescent solar prominences, cool condensations of plasma
in the solar corona that can last days to weeks
\citep{VialEngvold2015}.  In the case of the Sun, these condensations
fall back to the photosphere because gravity is stronger than any
magnetic tension or centrifugal force capable of sustaining them.
However, it has been understood at least since work by
\citet{Donati2000} that these ``prominence systems'' can be
longitudinally extended, forming a trapped ring of plasma.  While the
issue of why these previously known spectroscopic variables do not
show complex photometric variability remains open, plausible
explanations include that they are not in the required transiting
geometry, or that they lack the necessary source of opacity.  What we
have in the case of TIC\,141146667 is a system that finally exhibits
both sets of phenomena.

@@

The material's composition -- either pure plasma or dusty plasma --
can be clarified by time-series infrared spectrophotometry.  While
observations of CPVs in the optical suggest chromaticity consistent
with dust \citep{Tanimoto2020,Gunther2022,Koen2023}, electron
scattering in a plasma transiting over a spotted star might
also produce chromatic features \citep{Rackham2018}.  This
degeneracy is alleviated in the infrared, where the composition and
size distribution of any dust that is present could be determined by
measuring the extinction curve for a sample of CPVs from 1-20\,$\mu$m.
Composition and size distributions similar to debris from rocky bodies
seen around white dwarfs \citep{Reach2009} would suggest an extrinsic
origin channel.  Compositions and sizes similar to the interstellar
medium would suggest that dust can condense out from M dwarf winds,
similar to processes that occur around evolved stars
\citep{Marigo2008}.

Future models could take a range of steps to clarify the origin
of CPVs.  A first step for magnetohydrodynamic models that can
currently explain the H$\alpha$ features \citep[e.g.][]{Waugh2022}
would be to incorporate an opacity source that can reproduce the
photometric dips.  This effort will also require
exploring what stellar magnetic field topologies reproduce the
observed light curve shapes.  While previous
models in the rigid-field approximation have explored both dipolar
\citep{Townsend2008} and multipolar fields \citep{Krticka2022}, dynamo
simulations of fully-convective M dwarfs have suggested that
global-scale mean fields might be confined to a single hemisphere
\citep{Brown2020}; such fields would yield accumulation surfaces 
different from those that have been explored.  One related challenge
is that CPVs currently suffer a dearth of stellar surface field
measurements, unlike benchmark systems such as AB Dor and V374~Peg
\citep{Jardine2002,Vidotto2011}. This might be alleviated by
future spectropolarimetry, or by radio observations
sensitive to the magnetic field structure around the star
\citep[][]{Callingham2021,Kaur2024}.  Finally, the question of what
microphysical process yields the opaque material also merits future
theoretical attention. While recent advances can produce the
catastrophic cooling that leads to dense, cool clumps embedded in a
hot corona \citep{Daley-Yates2023,Daley-Yates2024}, a self-consistent
treatment of radiation propagating through such clumps has yet to be
performed. Whether dust grains can form in these plasma clumps
similarly deserves attention.
